CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report

### 2011-09-08 report/report.tex
Added parallelization graph
First draft complete

%----------------------------------------
% Parallelization
%----------------------------------------
\subsection{Parallelization}
\label{parallelization}

\begin{figure}[p]
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{paralleldeploymentmodel}
  \end{center}
  \caption{Deployment model of parallel indexer} 
  \label{fig:paralleldeploymentmodel}
\end{figure}

@@

One of the primary goals of this research project was to leverage the
effect of parallelization in reducing the time to index a document
collection. Due to time constraints, this was not done but the
resulting implementation lays a strong foundation for further research
in this area. Other avenues of interest include, but are not limited
to:

@@

>>>>

%----------------------------------------
% Parallelization
%----------------------------------------
\subsection{Parallelization}
\label{parallelization}
Given the modular architecture of our indexer, it is possible to
deploy multiple instances of the program to work concurrently on the same
indexing task. The instances interact via the database which
synchronizes their operations on the corpus and the index itself. To
investigate the effect of parallelization on the indexer operation, the
indexer was deployed as follows:

\begin{itemize}
\item Single database server storing documents, keywords, and the
  resulting index.
\item Four indexer nodes each running a single instance of the indexer
  program. 
\end{itemize}

This deployment as well as a description of the sequence of operation
are shown in figure \ref{fig:paralleldeploymentmodel}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{paralleldeploymentmodel}
  \end{center}
  \caption{Deployment model of parallel indexer} 
  \label{fig:paralleldeploymentmodel}
\end{figure}

In this experiment a corpus of 100,00 documents and 1000 keywords were
used, each document 5000 words in length. Tests were run using various
node combinations and the time taken to index the corpus was measured and
normalized to a throughput metric (documents/second) for convenient
comparison. Test results using the same number of nodes were averaged
to account for differences in the physical capabilities and
configuration of the worker nodes. Next, a series of tests was
performed using a corpus of 800,000 documents to investigate the
effect of corpus size on performance. The results of this experiment
are shown in figure \ref{fig:parallelization} and described in detail
in the following section.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{parallelization}
  \end{center}
  \caption{Performance characteristics of parallel indexer
    deployment. The \textit{ideal} curve shows the theoretical
    performance curve based on linear combination of average
    individual node performance over the corpus} 
  \label{fig:parallelization}
\end{figure}

\subsubsection{Results}
For the 100,000 document corpus, increasing the number of nodes
increases throughput but with diminishing returns. As the database
connection pool is saturated, this becomes the limiting factor of the
architecture. If the number of nodes were increased further, one would
expect the curve to plateau at some point. After this point, adding
additional nodes would offer no improvement and may even decrease
performance as the worker processes spend an increasing amount of time
context switching and contending for locks. These factors could be
mitigated up to a point by increasing batch sizes and tuning the
database. However, the characteristics of the curve are ultimately
unavoidable in the current architecture.

For the 800,000 document corpus, the performance curve shows an even
more severe degradation than the smaller corpus. Since we normalize
the performance metrics for the size of the document collection, one
would expect this curve to be more similar to that of the 100,000
document collection. A possible explanation for the deviation is that
for the larger collection, more documents were present in the
database and the resulting index larger. This results in longer table
scan times to retrieve and update records. This effect could be
alleviated through the use of indices and other database optimization
techniques, however, as for the smaller document
collection, this degradation is ultimately unavoidable.

@@

One of the primary goals of this research project was to leverage the
effect of parallelization in reducing the time to index a document
collection. We successfully demonstrated that performance gains could
be achieved for the indexing task through parallelization as well as
the fact that our architecture offers diminishing returns as the
number of nodes is increased and as the corpus becomes larger.

 Other avenues of interest include, but are not limited
to:

@@

      \item Conduct further parallelization experiments to determine
        the performance characteristics of our indexer over a wider
        range of operating conditions.
